---
title: "Packages"
output:
  distill::distill_article
---

<style>
d-title{
    padding: 2rem 1 1.5rem;
}
d-article {
    padding-top: 0em;
    font-size: 1em;
}
d-title h1 {
  margin-bottom: 0em;
  font-size: 2.25em;
}
d-article h3 {
  margin-top: 1em; 
  margin-bottom: 1em;
}
d-article h4 {
  margin-top: 0em; 
  margin-bottom: 0em;
}
d-article li {
    margin-bottom: 0.15em;
    margin-left: 0;
    padding-left: 0;
}
.visualization {
  text-align: center;
  margin: 0 0em;
}
</style>
### 1. **BASTION**: Bayesian Time-Series Decomposition
```{r datasim, echo = FALSE}
# arguments for figures:
cex_s = 0.8
cex.axis_s = 1
cex.axis_s_y = 0.9
cex.axis_s_x = 0.9
cex.main_s = 1.75
cex.lab_s = 1.25
lwd_s = 2.5
line = 2.5
cex_l = 0.8
margin_size1 = c(4,4,1,1)
generate_T <- function(n) {
  
  # Generate random parameters
  m <- runif(4, -50, 50)  # Slopes
  c <- runif(4, -25, 25)  # Intercepts
  b <- runif(3, 30, n%/%4)  # Breakpoints b1, b2, b3
  
  # Define breakpoints
  breakpoints <- cumsum(c(0, b, n - sum(b)))
  
  # Define the piecewise function
  T <- function(t) {
    if (t >= 0 && t < breakpoints[2]) {
      return(m[1] * 0.01 * (t - breakpoints[1]) + c[1])
    } else if (t >= breakpoints[2] && t < breakpoints[3]) {
      return(m[2] * 0.01 * (t - breakpoints[2]) + c[2])
    } else if (t >= breakpoints[3] && t < breakpoints[4]) {
      return(m[3] * 0.01 * (t - breakpoints[3]) + c[3])
    } else if (t >= breakpoints[4] && t <= n) {
      return(m[4] * 0.01 * (t - breakpoints[4]) + c[4])
    } else {
      return(NA)  # Out of range
    }
  }
  
  # Generate a sequence of t values
  t_values <- 1:n
  T_values <- sapply(t_values, T)
  
  # Return the results
  return(list(
    T_function = T,
    t_values = t_values,
    T_values = T_values,
    parameters = list(m = m, c = c, b = b, breakpoints = breakpoints)
  ))
}
gen_sim <- function(n,seed){
  #setting seed
  set.seed(seed)
  
  # trend componenet
  trend = generate_T(n)$T_values
  
  # remainder
  error_sd <- stochvol::svsim(n, mu = 4, phi = 0.98, sigma = 0.3, nu = Inf, rho = 0)
  errors = rnorm(n,0,error_sd$vol)
  vol = error_sd$vol
  # errors = rnorm(n,0,5)
  # vol = 5
  period1 <- 7
  period2 <- 30

  # Generate seasonality components
  # First seasonality term
  season1 <- rnorm(1,0,10)*sin(2 * pi * (1:n) / period1) + rnorm(1,0,10)*cos(2*pi*(1:n)/period1)
  season1 <- season1 - mean(season1)  # Adjust to make the first term 0
  
  #second seasonality term
  season2 <- rnorm(1,0,15)*sin(2 * pi * (1:n) / period2) + rnorm(1,0,15)*cos(2 * pi * (1:n) / period2) 
  season2 <- season2 - mean(season2) # Adjust to make the first term 0
  season = season1 + season2
  #Outlier term  
  n_outliers = rpois(1,5)
  outlier_ind = sample(c(5:n),size = n_outliers,replace = FALSE)
  outliers = numeric(n)
  outliers[outlier_ind] = rcauchy(n_outliers,loc = 35,scale = 4)*
    sample(c(-1,1),size = n_outliers,replace = TRUE)
  
  y = trend +season1 + season2 + outliers + errors
  return(as.data.frame(cbind(y,trend,season,season1,season2,outliers,errors,vol)))
}
y=gen_sim(500,2)
```

<div class="visualization">
```{r figure simulated data, echo = FALSE,fig.width = 8,fig.height = 2.75}
par(mar = margin_size1)
plot(y[,"y"], 
     type = "p",
     xlab = "",
     ylab = "",
     main = "",
     cex = cex_s, 
     cex.axis = cex.axis_s, 
     cex.main = cex.main_s,
     axes = FALSE)
axis(2,cex.axis = cex.axis_s_y)
axis(1,cex.axis = cex.axis_s_x)
abline(v = which(y[,c("outliers")] != 0),col = "red",lty = 2,lwd = lwd_s*3/5)
lines(y[,"trend"]+y[,"season"], type = "l", lwd = lwd_s, col = "blue")
lines(y[,"trend"], type = "l", lwd = lwd_s, col = "black")
legend("topright", lwd = 2,lty = c(1,1,2),col = c("black","blue","red"), 
       legend = c("Trend","Trend+Seasonality","Outliers"),
       cex = cex_l)
title(ylab = "Observations", cex.lab = cex.lab_s, line = line)
title(xlab = "Time", cex.lab = cex.lab_s, line = line)
```
</div>
  - **GitHub**: [Repository Link](https://jasoncho0914.github.io/BASTION/index.html)
  - **Non-Technical Guide**: [Guide Link](https://jasoncho0914.github.io/BASTION/articles/non-technical_guide.html)
  - **Preprint**: [Docs Link](paper/BASTION_20250124.pdf)
